---
title: "Feature Selection w/ Caret"
output: html_notebook
---

# Feature Selection with the Caret R Package

- How to remove redundant features from dataset
- How to rank features
- How to select features using recursive feature elimination

https://machinelearningmastery.com/feature-selection-with-the-caret-r-package/

```{r}
if(!require(mlbench)){install.packages("mlbench"); require(mlbench)} # common datasets to use
if(!require(caret)){install.packages("caret", dependencies = c("Depends", "Suggests")); require(caret)} # ML package and its dependencies. This will take awhile!
set.seed(370)
```

We will look at data relating to diabetes among Pima Indians. Notice how the outcome variable is categorical. This makes this a *classification* problem. When the outcome variable is continuous, it is a *regression* problem.
```{r}
data("PimaIndiansDiabetes")
?PimaIndiansDiabetes
head(PimaIndiansDiabetes)
```


Removing redundant features (absolute correlation cutoff)
```{r}
corr_matrix <- cor(PimaIndiansDiabetes[,1:8]) # correlations between all predictor vars
corr_matrix

cutoff <- 0.5 # should be higher in practice

highly_corr <- findCorrelation(corr_matrix, cutoff=cutoff)
print(colnames(PimaIndiansDiabetes)[highly_corr]) # age is highly correalted with pregnant
```

Ranking features by importance
```{r}
control <- trainControl(method="repeatedcv", number = 10, repeats = 3)

model <- train(diabetes ~., data=PimaIndiansDiabetes, method = "knn", preProcess = "scale", trControl = control)

importance <- varImp(model)

ggplot(importance)
```

Automatic Feature Selection

Uses different subsets of the predictor variables to build many models to identify variables that can be removed without compromising the accuracy model in a major way.
```{r}
# control using random forest
control <- rfeControl(functions = rfFuncs, method="cv", number=10)
results <- rfe(PimaIndiansDiabetes[,1:8], PimaIndiansDiabetes[,9], sizes = c(1:8), rfeControl = control) # this will take AWHILE...

results
ggplot(results)

# chosen features
predictors(results)
```


## More on caret
*caret* package (short for classification and regression training) contains functions to streamline the model training process for regression (output is continuous) and classification (output is categorical) problems. The package utilizes a number of R packges and all of them can be installed with the following command:

`install.packages("caret", dependencies = c("Depends", "Suggests"))`

Helpful Links:
- [Online book](http://topepo.github.io/caret/index.html)
- [Main help page](http://caret.r-forge.r-project.org/)


