---
title: "Fixing Plots"
output: html_notebook
---

# Making plots better

Good plots reflect *expressiveness* (showing data truthfully) and *effectiveness* (through encodings which people can interpret)

```{r}
if (!require(ggplot2)) { install.packages('ggplot2'); require(ggplot2) }
if (!require(dplyr)) { install.packages('dplyr'); require(dplyr) }
if (!require(data.table)) { install.packages('data.table'); require(data.table) }
if (!require(plyr)) { install.packages('plyr'); require(plyr) }

# loading some data...
data(mtcars)

View(mtcars)
```


# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars, aes(x=cyl)) + geom_bar(fill='darkgoldenrod2') +
  theme(panel.background = element_rect(fill = 'steelblue'),
  panel.grid.major = element_line(colour = "firebrick", size=3),
  panel.grid.minor = element_line(colour = "blue", size=1))
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}

freq_cars_by_cyl <- as.data.frame(table(mtcars %>% select(cyl))) # freq table of # cars by cylinder
freq_cars_by_cyl <- rename(freq_cars_by_cyl, c("Var1"="cyl")) # renaming a column

View(freq_cars_by_cyl)

ggplot(as.data.frame(freq_cars_by_cyl),
       aes(x = "", y = Freq, fill=cyl)) +
  geom_bar(width=1, stat = 'identity')  + # create bar plot
  coord_polar(theta="y") # turn to pie chart

```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars, aes(x=disp, y=hp)) + 
    geom_point(
        color="pink",
        fill="blue",
        shape=21,
        alpha=0.5,
        size=50,
        stroke = 4
        )
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars) +
  geom_point(aes(wt, hp, color=cyl)) +
  scale_colour_gradient(low = "red", high="green")
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
_hint_: check out facet_grid
```{r}
head(iris)

ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + geom_point(shape=4, size=20)
```

