---
title: "Building a Basic Model"
output: html_notebook
---
INFO 370A
Benjamin Xie & Greg Nelson
University of Washington

Adapted from Ch 24.3 of (R for Data Science)[http://r4ds.had.co.nz/model-building.html#what-affects-the-number-of-daily-flights], Garrett Grolemund & Hadley Wickham


Loading packages and data. See help documentation for more on the data.
```{r}
if(!require(tidyverse)){install.packages("tidyverse"); library(tidyverse)} 
if(!require(modelr)){install.packages("modelr"); library(modelr)} 
if(!require(lubridate)){install.packages("lubridate"); library(lubridate)} 
if(!require(nycflights13)){install.packages("nycflights13"); require(nycflights13)}

?flights
head(flights)
```

Data Cleaning
```{r}
daily <- flights %>% 
  mutate(date = make_date(year, month, day)) %>%
  group_by(date) %>%
  summarize(n = n())

# number of flights by day
ggplot(daily, aes(date, n)) + 
  geom_line() + 
  labs(title = "Number of flights out of NYC by day")
```

Controlling by day of week
```{r}
daily <- daily %>%
  mutate(wday = wday(date, label=TRUE))

ggplot(daily, aes(wday, n)) + 
  geom_boxplot()

ggplot(daily, aes(n)) +
  geom_histogram() + 
  facet_wrap(~wday) +
  labs(title = "Number of flights by day of week")
```

Using model to remove pattern of flights by day of week. 
*TODO*: try out a nonlinear model and see what happens!
```{r}
# first fit model and display predictions over original data
mod <- lm(n ~ wday, data = daily)

grid <- daily %>%
  data_grid(wday) %>% 
  add_predictions(mod, "n")

# showing prediction
ggplot(daily, aes(n)) + 
  geom_histogram() + 
  facet_wrap(~wday) + 
  geom_vline(data = grid, aes(xintercept = n), color = "red")

# now compute and visualize residuals
daily <- daily %>%
  add_residuals(mod)

ggplot(daily, aes(date, resid)) + 
  geom_ref_line(h=6) + 
  geom_line()

daily2 <- daily

# given residuals, maybe we want to remove odd days (Mother's Day, July 4th)
daily2 <- daily2 %>%
  filter(resid > -100)
```

Looking at effect of time of year
```{r}
if(!require(splines)){install.packages("splines"); library(splines)} 
if(!require(MASS)){install.packages("MASS"); require(MASS)}
```

Using robust fitting of linear modles (RLM) to create a model with interactions between day of week and time of year
```{r}
mod <- MASS::rlm(n ~ wday * ns(date, df=5), data = daily)  # degrees of freedom was arbitrary choice

# plot of models
daily2 %>%
  data_grid(wday, date = seq_range(date, n=13)) %>%
  add_predictions(mod) %>%
  ggplot(aes(date, pred, color=wday)) + 
    geom_line()+
    geom_point()

# now compute and visualize residuals
daily2 <- daily2 %>%
  add_residuals(mod)

ggplot(daily2, aes(date, resid)) + 
  geom_ref_line(h=6) + 
  geom_line()
```