---
title: "Class 6.3: Building Basic Models"
output: html_notebook
---
INFO 370A
Benjamin Xie & Greg Nelson
University of Washington


Adapted from Ch 23, 24 of (R for Data Science)[http://r4ds.had.co.nz], Garrett Grolemund & Hadley Wickham

```{r}
if(!require(tidyverse)){install.packages("tidyverse"); library(tidyverse)} 
if(!require(modelr)){install.packages("modelr"); library(modelr)} 
if(!require(splines)){install.packages("splines"); require(splines)}
```

# Interaction Effects
Interactions (continuous & categorical)

`sim3`: categorical predictor (x2) and continuous predictor (x1)
```{r}
head(sim3)

ggplot(sim3, aes(x1, y)) + 
  geom_point(aes(color = x2))
```

2 possible models for `sim3`
- `+` makes it so model estimates each effect *independent* of all others
- `*` model can fit interaction
```{r}
mod1 <- lm(y ~ x1 + x2, 
           data = sim3)
mod2 <- lm(y ~ x1 * x2, 
           data = sim3) 

# y = a_0 + a_1 * x1 + a_2 * x2 + a_12 * x1 * x2

```

To visualize these models, 2 new tricks:
- pass both `x1` and `x2` for `data_grid()`
- use `gather_predictions()` to generate predictions from both models simultaneously, adding prediction as a row
  - complement is `spread_predictions()`, which adds each prediction to a new column
```{r}
grid <- sim3 %>%
  data_grid(x1, x2) %>%
  gather_predictions(mod1, mod2)

grid

# data_sim3 <- sim3 %>%
#   mutate(
#     mod1 = mad_fit_intercept + mad_fit_slope * data_sim1$x
#   )

ggplot(sim3, aes(x1, y, color = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) +
  facet_wrap(~model)
```

Judging model using residuals
- residuals in mod1 show pattern (which is bad)
```{r}
sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

head(sim3)

ggplot(sim3, aes(x1, resid, color = x2)) + 
  geom_point() +
  facet_grid(model ~ x2)
```


# Looking at non-linear data

Generating some non-linear data
```{r}
sim5 <- tibble(
  x = seq(0, 3.5 * pi, length = 50),
  y = 4 * sin(x) + rnorm(length(x))*2
)

# plotting random points
ggplot(sim5, aes(x, y)) +
  geom_point()
```


```{r}
# creating a bunch of "polynomial" models with various degrees of freedom
mod1 <- lm(y ~ ns(x, df=1), data=sim5)
mod2 <- lm(y ~ ns(x, df=2), data=sim5)
mod3 <- lm(y ~ ns(x, df=3), data=sim5)
mod4 <- lm(y ~ ns(x, df=4), data=sim5)
mod5 <- lm(y ~ ns(x, df=5), data=sim5)
mod20 <- lm(y ~ ns(x, df=20), data=sim5)

grid <- sim5 %>%
  data_grid(x = seq_range(x, n= 50, expand = 0.1)) %>%
  gather_predictions(mod1, mod2, mod3, mod4, mod5, mod20, .pred = "y")

# plotting predictions
ggplot(sim5, aes(x,y)) +
  geom_point() + 
  geom_line(data = grid, color = "red") + 
  facet_wrap(~ model)
```

Looking at residuals
```{r}
#calculating and plotting residuals
sim5 %>%
  gather_residuals(mod1, mod2, mod3, mod4, mod5, mod20) %>%
  ggplot(aes(x, resid)) + 
  geom_point() + 
  facet_wrap(~model)
```


# Factors affecting number of flights

Loading packages and data. See help documentation for more on the data.
```{r}
if(!require(lubridate)){install.packages("lubridate"); library(lubridate)} 
if(!require(nycflights13)){install.packages("nycflights13"); require(nycflights13)}

?flights
head(flights)
```

Data Cleaning
```{r}
daily <- flights %>% 
  mutate(date = make_date(year, month, day)) %>%
  group_by(date) %>%
  summarize(n = n())

# number of flights by day
ggplot(daily, aes(date, n)) + 
  geom_line() + 
  geom_point() +
  labs(title = "Number of flights out of NYC by day")
```

Controlling by day of week
```{r}
daily <- daily %>%
  mutate(wday = wday(date, label=TRUE))

ggplot(daily, aes(wday, n)) + 
  geom_boxplot()

ggplot(daily, aes(n)) +
  geom_histogram() + 
  facet_wrap(~wday) +
  labs(title = "Number of flights by day of week")
```

Using model to remove pattern of flights by day of week. 
```{r}
# first fit model and display predictions over original data
mod <- lm(n ~ wday, data = daily)

grid <- daily %>%
  data_grid(wday) %>% 
  add_predictions(mod, "n")

# showing prediction
ggplot(daily, aes(n)) + 
  geom_histogram() + 
  facet_wrap(~wday) + 
  geom_vline(data = grid, aes(xintercept = n), color = "red")

# now compute and visualize residuals
daily <- daily %>%
  add_residuals(mod)

ggplot(daily, aes(date, resid)) + 
  geom_ref_line(h=6) + 
  geom_line() + 
  labs(title = "Residuals after linear model for day of week")

daily2 <- daily

# given residuals, maybe we want to remove odd days (Mother's Day, July 4th)
# daily2 <- daily2 %>%
  # filter(resid > -100)
```

Looking at effect of time of year
```{r}
if(!require(splines)){install.packages("splines"); library(splines)} 
if(!require(MASS)){install.packages("MASS"); require(MASS)}
```

Using robust fitting of linear modles (RLM) to create a model with interactions between day of week and time of year
```{r}
mod <- MASS::rlm(n ~ wday * ns(date, df=5), data = daily)  # degrees of freedom was arbitrary choice

# plot of models
daily2 %>%
  data_grid(wday, date = seq_range(date, n=13)) %>%
  add_predictions(mod) %>%
  ggplot(aes(date, pred, color=wday)) + 
    geom_line()+
    geom_point()

# now compute and visualize residuals
daily2 <- daily2 %>%
  add_residuals(mod)

ggplot(daily2, aes(date, resid)) + 
  geom_ref_line(h=6) + 
  geom_line()
```